---
title: "R Packages"
author: "Rob Harbert"
date: "5/17/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Packages Book

http://r-pkgs.had.co.nz/intro.html

## Useful R libraries for package writing and management

```{r, eval = FALSE}
library(devtools)
require(roxygen2)
```

### Build package skeleton

```{r, eval = FALSE}
rm(list = ls()) # clear workspace before building

getwd()
dir.create('./newR/')
setwd(paste(getwd(), '/newR/', sep=''))
create('./')

roxygenize('./') #Builds description file and documentation


```

roxygen2 is great, BUT we have to write function documentation in a specific way. Write the next block to a file and put it in the R subdirectory.

```{r, eval=FALSE}
#' Echo
#' 
#' This function echos whatever you give it.
#' 
#' @param echo A word or sentence to echo
#'
#' @export
#' @examples
#'
#' echo('This is a test')


echo = function(echo){
  return(echo)
}
    
    
```

Then roxygenize

```{r, eval = FALSE}
roxygenize('./')
```

Or just check with devtools

```{r, eval = FALSE}
check(cran=TRUE)
```

Once you check() the library this can be installed as usual:

```{r, eval = FALSE}
library(newR)
```

AND we can upload that repository to GitHub and install from there. I advise using the GitHub desktop app: https://desktop.github.com/ (Mac/Windows). GitKraken might be OK too: https://www.gitkraken.com/download/mac 

```{r, eval=FALSE}
install_github('rsh249/newR')
```

Git install

```{bash, eval=FALSE}
#set up git
git config --global user.name 'Your Name'
git config --global user.email 'your@email.com'
git config --global credential.helper osxkeychain  #on mac only

#local git repository in library folder
cd newR
git init
git add
git commit

#then add to github
git remote add origin https://github.com/username/new_repo
git push -u origin master
 
#Then for updates:
git add .
git commit -m 'Add files'
```


