---
title: "R Basics Lab"
author: "Rob Harbert"
date: "5/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Basics

## Outline:

* Setting up on Lyell
* Interactive R - Data I/O refresher, calling functions, libraries, data visualization.
* Vectorizing
* Scripting R
* Defining R functions

## Lyell

We will be working on a teaching cluster maintained by the SICG. You will get printed instructions on how to log in. We have pre-installed most of the R libraries we will need for this course, but where possible I will include installation details for libraries we are using here.

## Interactive R

R is very handy for it's interactive command line interface. To get started type: 'R' at your command line.

What version of R do you have?


```{r fig.width=7, fig.align='center', echo=FALSE}
library(png)
library(grid)
img <- readPNG("./docs/Rprompt.png")
 grid.raster(img)
```
Or
```{r}
version
```

We can easily get started with the R command promp open.

```{r}
x=2
print(x) ##Print method
class(x)

x=seq(1:10) # Create a vector
class(x)
print(x)
print(x[1]) # First index of vector
print(x[1:5])

y = matrix(nrow=5, ncol=5) # create a 5x5 matrix
print(y)
class(y)
y[1,1] = 5
print(y)
y[,1]  = x[1:5]
print(y)
class(y[,1])

y = cbind(seq(1:5), 
          seq(1:5),
          seq(1:5),
          seq(1:5),
          seq(1:5))

class(y)


```

## Data I/O

Read table/tab/csv text files:
read.table()
read.csv()
read.delim()

```{r}
cars = read.table('../data/mtcars.csv', header=T, sep = ',') # Read a comma separated values file
head(cars)

cars = read.csv('../data/mtcars.csv')
cars = read.csv2('../data/mtcars.csv') ## Interesting behavior here, will be somewhat faster

cars = read.delim('../data/mtcars.csv', sep=',')


```

fread - Faster read

```
library(data.table)
ebd = fread('../data/ebd_trim2.csv', sep = '\t')
```


## ggplot

### ggmaps

## Loops

## apply family functions

## Writing Functions





